---
title: "ANA 515 Assignment 4 Data Analytics Project"
author: "Shreyesh Shetty"
date: "2022-10-14"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 90
---

#1 Discuss the business problem/goal The main goal is to build a recommendation engine to
recommend movies to the user.When we watch movies on any OTT platform we start getting
suggestions for similar movies, the problem is to identify how this recommendation engine
works and how the system learns from watching patterns to provide relevant suggestions.

#2. identify where the dataset was retrieved The data is retrieved from the Movie Lens
dataset which has 2 datasets that contains movie names and Movie genre and 2nd data set
has the ratings of those movies.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tinytex::install_tinytex(force=TRUE)
library(recommenderlab)
library(ggplot2)                       
library(data.table)
library(reshape2)
#3.	identify the code that imported and saved your dataset in R 

setwd("C:/Users/Shreyesh Shetty/Documents/Data Storage")                       
movie_data <- read.csv("movies.csv",stringsAsFactors=FALSE)
rating_data <- read.csv("ratings.csv")
str(movie_data)

summary(movie_data)

head(movie_data)

summary(rating_data)

head(rating_data)
```

#4. describe your data set (using the common attributes such as #rows, #columns, variable
names, types, means, SD, #min/max, NAs, etc...)

The movie data set has `r nrow(movie_data)` rows and `r ncol(movie_data)` columns, mean is
`r mean(movie_data)` . Standard deviation `r sd(movie_data$movieId)`. Minimum is
`r min(movie_data)`. maximum is `r max(movie_datat)`.

The rating data set has `r nrow(rating_data)` rows and `r ncol(rating_data)` , mean of
userId is `r mean(rating_data$userId)`, movieId is `r mean(rating_data$movieId)`,rating is
`r mean(rating_data$rating)` and Timestamp is `r mean(rating_data$timestamp)`.

Standard deviation of userId is `r sd(rating_data$userId)`and movieId is
`r sd(rating_data$movieId)`,rating is `r sd(rating_data$rating)` and Timestamp is
`r sd(rating_data$timestamp)`.

Minimum of userId is `r min(rating_data$userId)`and movieId is
`r min(rating_data$movieId)`,rating is `r min(rating_data$rating)` and Timestamp is
`r min(rating_data$timestamp)`.

maximum userId is `r min(rating_data$userId)`and movieId is
`r min(rating_data$movieId)`,rating is `r min(rating_data$rating)` and Timestamp is
`r min(rating_data$timestamp)`.

#5. discuss any data preparation, missing values and errors The dataset is pretty much
clean, but not in terms of the business problem we are trying to figure out we will have
to convert genres present in the movie_data dataframe into a more usable format for which
we will first we will first create a one-hot encoding to create a matrix that comprises of
corresponding genres for each of the films.by matrix i mean, the column genre has a
different genres which are in one column, which needs to be split into different columns.
Then need to convert the characters to integer. Next we will create a 'search matrix' that
will allow us to perform an easy search of the films based on genre present in our list.
We will bind the new dataframes created into one to give a value to each of the movies
according to their genre. Next we will have to convert our matrix into a sparse matrix
using dcast. Explore similar data and then conduct data preparation in the following three
steps --Selecting useful data, Normalizing data,Binarizing the data. Some users will give
very high or very low ratings to all the watched movies, this will act as a bias while
implementing model. Standardize the data by normalizing the numerical values to a common
scale value without distorting the range of values. Convert the data to binary data the
data that will have 2 discrete values i.e.1&0, define a matrix that will consist of 1 if
the rating is above 3 and otherwise it will be 0

```{r ,results='asis'}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors=FALSE)
library(data.table)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE) 
colnames(movie_genre2) <- c(1:10)
list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre
for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col]) 
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)

SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix) 

ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix

recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)

lapply(recommendation_model, "[[", "description")

recommendation_model$IBCF_realRatingMatrix$parameters

similarity_mat <- similarity(ratingMatrix[1:4, ],
                               method = "cosine",
                               which = "users")
as.matrix(similarity_mat)
image(as.matrix(similarity_mat), main = "User's Similarities")

movie_similarity <- similarity(ratingMatrix[, 1:4], method =
                                 "cosine", which = "items")
as.matrix(movie_similarity)
image(as.matrix(movie_similarity), main = "Movies similarity")

rating_values <- as.vector(ratingMatrix@data)
unique(rating_values) # extracting unique ratings

Table_of_Ratings <- table(rating_values) # creating a count of movie ratings
Table_of_Ratings

library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie_data,
                                         movie_data$movieId == table_views[index,1])$title)
}
table_views[1:6,]

ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films")
   

image(ratingMatrix[1:20, 1:25], axes = FALSE, main = "Heatmap of the first 25 rows and 25 columns")

movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                             colCounts(ratingMatrix) > 50]
movie_ratings

minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")

average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("steelblue"), col=I("red")) +
  ggtitle("Distribution of the average rating per user")


normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)
image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                          colCounts(normalized_ratings) > minimum_users],
main = "Normalized Ratings of the Top Users")

binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.95)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.95)
#movies_watched <- binarize(movie_ratings, minRating = 1)
good_rated_films <- binarize(movie_ratings, minRating = 3)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
colCounts(movie_ratings) > binary_minimum_users],
main = "Heatmap of the top users and movies")

sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```

#6. discuss the modeling The algorithm will now identify the k most similar items and
store their number for which we will use cosine method Using the getModel() function, we
will retrieve the recommen_model. We will then find the class and dimensions of our
similarity matrix that is contained within model_info. Finally, we will generate a
heatmap, that will contain the top 20 items and visualize the similarity shared between
them. we will carry out the sum of rows and columns with the similarity of the objects
above 0. We will visualize the sum of columns through a distribution as follows We will
create a top_recommendations variable which will be initialized to 10, specifying the
number of films to each user. We will then use the predict() function that will identify
similar items and will rank them appropriately. Here, each rating is used as a weight.
Each weight is multiplied with related similarities. Finally, everything is added in the
end.

#7. produce and discuss the output The output is being presented in the movie_user2 by
using each rating as a weight. Each weight is then multiplied with related similarities.
Finally, everything is added in the end.

#8. provide explanation with any visuals The final visualization displays a graphical
representation of the multiplied value of related similarities with each weight for each
of the recommendation. Also we could have used a different kind distribution to display
how movies that fall under more genres have high or low rating and can be used for
someother kind of predictive model to make the movies more interesting.

```{r,results='asis'}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters

recommen_model <- Recommender(data = training_data,
                          method = "IBCF",
                          parameter = list(k = 30))
recommen_model
class(recommen_model)

model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
   main = "Heatmap of the first rows and columns")

sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)
sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red"))+ ggtitle("Distribution of the column count")

top_recommendations <- 10 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                          newdata = testing_data,
                          n = top_recommendations)
predicted_recommendations

user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie_data,
                                         movie_data$movieId == movies_user1[index])$title)
}
movies_user2

recommendation_matrix <- sapply(predicted_recommendations@items,
                      function(x){ as.integer(colnames(movie_ratings)[x]) }) # matrix with the recommendations for each user
#dim(recc_matrix)
recommendation_matrix[,1:4]



```

```{r }

```

```{r}

```
